---
title: "Project"
output: html_document
date: "2024-04-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries}
library(tidyverse)
library(readxl)
library(magrittr)
library(stringi)
```

Important note: This is NOT a production model. This is an exploratory rmd file used before I had permission from Choisy to use his data, which does have a cleaner .rda file. Hence, the code herein are my numerous attempts to decipher a very messy .xlsx file.

```{r Summon Data}
roster = list() 
# Summons.
for (i in 1:5){
tbvn_years = readxl::read_excel("./data/TB data.xlsx",sheet = i) %>%
  # Remove any NA entries in Quarter, since these are likely extraneous rows.
  filter(!is.na(Quarter)) %>%
  mutate(Date = as.numeric(Quarter)) %>%
  filter(is.na(Date)) %>%
  mutate(Province = str_remove(Quarter,"[0-9]+") %>%
           ifelse(. == "", NA,.),
         across(2:10,~gsub("[^0-9.]", "",.)),
         across(2:10,~as.numeric(.)),
         temp = ifelse(grepl("^[0-9]\\.[0-9]*[1-9]",`AFB Positive New`),0,1)) %>%
  # Fills out the province column with its prior entries, which are VN's provinces.
  fill(Province) %>%
  # Discard original Quarter column.
  select(-c(Quarter)) %>%
  # Relocate province and date to the start.
  relocate(c("Province","Date"))
roster[[i]] = tbvn_years}
```

```{r Summon Data}
clean = function(number){readxl::read_excel("./data/TB data.xlsx",sheet = number) %>%
  # Remove any NA entries in Quarter, since these are likely extraneous rows.
  filter(!is.na(Quarter)) %>%
  mutate(Date = as.numeric(Quarter)) %>%
  filter(is.na(Date)) %>%
  select(-c(Date)) %>%
  mutate(across(2:10,~gsub("[^0-9.]", "",.)),
         across(2:10,~as.numeric(.)),
         across(2:10,~ifelse(grepl("^[0-9]*\\.[0-9]*[1-9]",.),. * 1000,.))) %>%
    select(c("Quarter","AFB Positive New"))}

tbvn_years_2010 = clean(1)
tbvn_years_2011 = clean(2)
tbvn_years_2012 = clean(3)
tbvn_years_2013 = clean(4)
tbvn_years_2014 = clean(5)
tbvn_years_2015 = clean(6)

final_2010 = tbvn_years_2010 %>%
  tail(n = 1)
```

```{r 2010}
readxl::read_excel("./data/TB data.xlsx",sheet = 1) %>%
  filter(Quarter != "Total") %>%
  # Create Province from quarter, discarding all numbers.
  # This isolates province names and total.
  # Then, convert into LATIN-ASCII.
  mutate(Province = str_remove(Quarter,"[0-9]+") %>%
           ifelse(. == "", NA,.) %>%
           stri_trans_general(.,"Latin-ASCII"),
         # Province is empty wherever quarter had a number.
         # ifelse empties the original province names and quarter.
         Quarter = as.factor(ifelse(is.na(Province),
                          Quarter,
                          NA))) %>%
  # Replicates province name each time.
  fill(Province) %>%
  # Keep if quarter is not empty, or Province is Total.
  filter(!is.na(Quarter)) %>%
  mutate(across(2:10,~gsub("[^0-9.]", "",.)),
         across(2:10,~as.numeric(.)),
         across(2:10,~ifelse(grepl("^[0-9]*\\.[0-9]*[1-9]",.),. * 1000,.))) %>%
  relocate(Province) %>%
  select(c(1:3)) %>%
  arrange(Province)
```

```{r Provinces}
provinces = read_csv("./data/verify_provinces.csv") %>%
  # Latinize province names, then discard province/city appendage. Trim.
  # Creates "Verify" column for comparison.
  mutate(Province = stri_trans_general(`Province/city`,"Latin-ASCII") %>%
           gsub("province", "", .) %>%
           gsub("city", "", .) %>%
           str_trim(),
         Verify = "Verify") %>%
  select(c("Province","Verify")) %>%
  filter(!is.na(Province))

temp %>% 
  # Title-case province.
  mutate(Province = str_to_title(Province)) %>%
  # Merge by shared provinces.
  merge(.,provinces, by = "Province", all=T) %>%
  # Quarter and Verify can only be NA if Province fails to match.
  # This allows validation for provinces that are misspelled 
  filter(is.na(Quarter) | is.na(Verify))
```

Ba Ria - Vung Tau, Bac Can, TP Ho Chi Minh, and Thua Thien Hue all exist, but are misspelled. Only TB Hospital seems entirely foreign. Since it can not be tested with a province, this implies that this can be discarded.
