---
title: "HW 1"
output:
  word_document: default
date: "2023-09-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Initialization}
CVD_cleaned = read.csv("./CVD_cleaned.csv",header = T)
library(magrittr)
library(tidyverse)
library(ggcorrplot)
library(dplyr)
```

# Below are categoricals.

```{r General Health, Bar Chart}
cus_order = c("Excellent","Very Good", "Good", "Fair", "Poor")

CVD_cleaned %>% 
  ggplot(aes(
    x = factor(General_Health, level = cus_order), 
    fill = General_Health)
    ) +
  geom_bar() +
  geom_text(
    stat='count', 
    aes(label=..count..), 
    vjust=1.5
    ) +
  scale_fill_manual(
    breaks = cus_order,
    values = c("Excellent" = "#0bff00", "Very Good" = "#55b100", "Good" = "#859a00", "Fair" = "#d6cb00","Poor" = "#ff0000")
    ) + 
  labs(x = "\n", 
       y = "Count", 
       fill = "General Health",
       title = "Respondents per health status")
```

```{r Exercise, Pie Chart}
CVD_cleaned %>% count(Exercise) %>% arrange(desc(Exercise)) %>%
  mutate(
    prop = round(n/sum(n)*100,1),
    lab.ypos = cumsum(prop)-.4*prop
    ) %>%
  mutate(label = paste0(Exercise,"\n", prop, "%")) %>%
  ggplot(aes(x="", y=prop, fill=Exercise)) +
  geom_bar(stat="identity", width=.5) +
  geom_text(
    aes(
      y = lab.ypos, 
      label=label),
    size = 5) +
  coord_polar("y", start=0, direction=-1) +
  theme_void() +
  theme(legend.position = "F") +
  scale_fill_manual(
    values = c("Yes" = "#55b100", "No" = "#dc1919")
  ) +
  labs(title = "Relative frequencies for exercise")
```

```{r Checkup, Relative Frequency}
CVD_cleaned_cus = CVD_cleaned %>% mutate(Checkup = recode(
    Checkup, 
    "Within the past year" = "Past year", 
    "Within the past 2 years" = "Past 2 years", 
    "Within the past 5 years" = "Past 5 years", 
    "5 or more years ago" = "5 or more years",
    "Never" = "Never")
    ) 
cus_order_checkup = c("Never","5 or more years","Past 5 years", "Past 2 years","Past year") 

CVD_cleaned_cus %>% 
  ggplot(aes(
    x = factor(Checkup,levels = cus_order_checkup), 
    y = after_stat(count/sum(count)),
    fill = factor(Checkup,levels = cus_order_checkup))
    ) +
  geom_bar() +
  scale_y_continuous(labels=scales::percent,
                     limits=c(0,1)) +
  geom_text(
    stat='count', 
    aes(label=paste(round(
      after_stat(count/sum(count)),3)*100,
      "%")),
    vjust=-0.5
    ) + 
  labs(x = "\n", 
       y = "Rate", 
       fill = "Checkup",
       title = "Respondents per Checkup type") +
  scale_fill_manual(
    values = c("Past year" = "#0bff00", "Past 2 years" = "#55b100", "Past 5 years" = "#859a00", "5 or more years" = "#d6cb00","Never" = "#ff0000")) +
  theme(axis.text.x=element_text(angle=-10))

```

```{r Heart Disease, Table}
table(CVD_cleaned$Heart_Disease)
```

```{r Skin Cancer, Bar Chart}
cus_order_yn = c("Yes","No") 
# Removes scientific notation.
options(scipen = 999)

CVD_cleaned %>% 
  ggplot(aes(
    x = factor(Skin_Cancer, level = cus_order_yn), 
    fill = Skin_Cancer)
    ) +
  geom_bar() +
  geom_text(
    stat='count', 
    aes(label=..count..), 
    vjust=-0.25
    ) +
  scale_fill_manual(
    breaks = cus_order_yn,
    values = c("No" = "#0bff00","Yes" = "#ff0000")
    ) + 
  labs(x = "Skin cancer response", 
       y = "Amount of patients",
       title = "Skin cancer presence", 
       fill = "Response") 
```

```{r Other Cancer, Pie Chart}
CVD_cleaned %>% count(Other_Cancer) %>% arrange(desc(Other_Cancer)) %>%
  mutate(
    prop = round(n/sum(n)*100,1),
    lab.ypos = cumsum(prop)-0.6*prop
    ) %>%
  mutate(label = paste0(Other_Cancer,"\n", prop, "%")) %>%
  ggplot(aes(x="", y=prop, fill=Other_Cancer)) +
  geom_bar(stat="identity", width=.5) +
  geom_text(
    aes(
      y = lab.ypos, 
      label=label),
    size = 4) +
  coord_polar("y", start=0, direction=-1) +
  theme_void() +
  theme(legend.position = "F") +
  scale_fill_manual(
    values = c("No" = "#55b100", "Yes" = "#dc1919")
  ) +
  labs(title = "Relative frequencies for other cancers")
```

```{r Depression, Table}
table(CVD_cleaned$Depression)
```

```{r Diabetes, Relative Frequency}
CVD_cleaned_cus = CVD_cleaned %>% mutate(Diabetes = recode(
    Diabetes, 
    "No" = "No", 
    "Yes" = "Yes", 
    "No, pre-diabetes or borderline diabetes" = "Pre/border diabetic", 
    "Yes, but female told only during pregnancy" = "Told yes during pregnancy")
    )

CVD_cleaned_cus %>% 
  ggplot(aes(
    x = Diabetes, 
    y = after_stat(count/sum(count)),
    fill = Diabetes)
    ) +
  geom_bar() +
  scale_y_continuous(labels=scales::percent,
                     limits=c(0,1)) +
  geom_text(
    stat='count', 
    aes(label=paste(round(
      after_stat(count/sum(count)),3)*100,
      "%")),
    vjust=-0.5
    ) + 
  labs(x = "\n", 
       y = "Rate", 
       fill = "Diabetes",
       title = "Respondents per diabetes type") +
  # These colors approximately resemble the grey-blue of the diabetes ribbon.
  scale_fill_manual(
    values = c("No" = "#c1c1c1", "Pre/border diabetic" = "#a0bce1", "Told yes during pregnancy" = "#80a8e2", "Yes" = "#518be1")
  ) +
  theme(axis.text.x=element_text(angle=-10))
```

```{r Arthritis, Bar Graph}
cus_order_yn = c("Yes","No") 
options(scipen = 999)

CVD_cleaned %>% 
  ggplot(aes(
    x = factor(Arthritis, level = cus_order_yn), 
    fill = Arthritis)
    ) +
  geom_bar() +
  geom_text(
    stat='count', 
    aes(label=..count..), 
    vjust=-0.25
    ) +
  scale_fill_manual(
    breaks = cus_order_yn,
    values = c("No" = "#0bff00","Yes" = "#ff0000")
    ) + 
  labs(x = "Arthritis response", 
       y = "Amount of patients",
       title = "Arthritis presence", 
       fill = "Response") 
```

```{r Sex, Pie Chart}
CVD_cleaned %>% count(Sex) %>% arrange(desc(Sex)) %>%
  mutate(
    prop = round(n/sum(n)*100,1),
    lab.ypos = cumsum(prop)-0.6*prop
    ) %>%
  mutate(label = paste0(Sex,"\n", prop, "%")) %>%
  ggplot(aes(x="", y=prop, fill=Sex)) +
  geom_bar(stat="identity", width=.5) +
  geom_text(
    aes(
      y = lab.ypos, 
      label=label),
    size = 4) +
  coord_polar("y", start=0, direction=-1) +
  theme_void() +
  theme(legend.position = "F") +
  scale_fill_manual(
    values = c("Male" = "#00A3E7", "Female" = "#FF6A66")
  ) +
  labs(title = "Relative frequencies for sex")
```

```{r Age Category, Bar Graph}
CVD_cleaned %>% 
  ggplot(aes(
    x = Age_Category, 
    fill = Age_Category)
    ) +
  geom_bar() +
  geom_text(
    stat='count', 
    aes(label=..count..), 
    vjust=-0.5,
    size = 3
    ) + 
  labs(x = "Categories", 
       y = "Patients per Category", 
       fill = "Ages",
       title = "Age categories")
```

```{r Smoking History, Bar Graph}
table(CVD_cleaned$Smoking_History)
```

## Below are quantitative variables.

```{r Correlation matrix of each quantitative}
matrix = CVD_cleaned %>% select_if(is.numeric) %>%
  drop_na() %>% round(digits = 2)
matrix = cor(matrix)
colnames(matrix) = c("Height","Weight","BMI","Alcohol Cons.","Fruit Cons.","Greens Cons.","Fried Potatoes Cons.")
rownames(matrix) = c("Height","Weight","BMI","Alcohol Cons.","Fruit Cons.","Greens Cons.","Fried Potatoes Cons.")
ggcorrplot(matrix, lab=T, lab_size=3,type = "upper")
```

# Because most of the correlations are weak, only BMI-Weight, Weight-Height, and Greens-Fruits have been charted.

```{r Fruit and Green}
# Main scatterplot with trend line. This will take a long time to load since there are 308k data points.
CVD_cleaned %>%
  ggplot(aes(x = Fruit_Consumption, y = Green_Vegetables_Consumption)) + 
  geom_point() +
  geom_smooth(method="lm", se = F) +
  labs(
    Title = "Fruit and Vegetable Consumption Scatterplot",
    x = "Fruit Consumption",
    y = "Green Vegetables Consumption"
)
```

```{r Weight and BMI}
CVD_cleaned %>%
  ggplot(aes(x = Weight_.kg., y = BMI)) + 
  geom_point() +
  geom_smooth(method="lm", se = F) +
  labs(
    Title = "Weight and BMI Scatterplot",
    x = "Weight",
    y = "BMI"
)
```


## Note: everything below this comment are experimental graphs. They were not explicitly requested, and I made them purely out of curiosity with the limits of R-based data visualization. Much of this code was based on a wide selection of online sources.

```{r General Health grouped by BMI, grouped bars}
# Below code creates a custom order for health, then creates BMI categories that include the lower but exclude the higher, allowing someone with a 24.99999 BMI to be classed into healthy. These are all stored inside of a CVD_cleaned_cus to avoid damaging the original dataframe.
cus_order_health = c("Excellent","Very Good", "Good", "Fair", "Poor")
CVD_cleaned_cus = CVD_cleaned %>% 
  mutate(BMI = cut(
    BMI, 
    breaks = c(0,18.5,25,30,35,Inf),
    include.lowest = TRUE,
    right = F,
    labels = c("Underweight", "Healthy", "Overweight", "Obese", "Morbidly Obese"))
    )

CVD_cleaned_cus %>% group_by(General_Health,BMI) %>%
  summarise(counts = n(),.groups = "keep"
    ) %>%
  ggplot(
    aes(
      x = factor(General_Health,level=cus_order_health),
      y = counts, 
      fill = BMI)
    ) +
  geom_bar(
    position="dodge", 
    stat = "identity"
    ) +
  geom_text(
    aes(label=counts),
    hjust = -0.1,
    size=2.5,
    position = position_dodge(0.9)
    ) +
  scale_fill_manual(
    values = c("Underweight" = "#808080", "Healthy" = "#0bff00", "Overweight" = "#859a00", "Obese" = "#d6cb00","Morbidly Obese" = "#ff0000")
    ) +
  coord_flip(ylim = c(0, 45000)) + 
  labs(
    x = "General Health", 
    y = "BMI Count per General Health \n",
    title = "Health, Split by BMI bins", 
    fill = "BMI Bins")
```

```{r General Health grouped by BMI, facet wrap view}
# A past version of this data visualization had a fixed scale for all five BMI classes, but this had issues with gleaning patterns from any given BMI class, which I would assume is more of value in medical sciences.
CVD_cleaned_cus = CVD_cleaned %>% 
  mutate(General_Health = recode(
    General_Health, 
    "Excellent" = "Exc", 
    "Very Good" = "V Good", 
    "Good" = "Good", 
    "Fair" = "Fair", 
    "Poor" = "Poor")
    ) %>%
    mutate(BMI = cut(
    BMI, 
    breaks = c(0,18.5,25,30,35,Inf),
    include.lowest = TRUE,
    right = F,
    labels = c("Underweight", "Healthy", "Overweight", "Obese", "Morbidly Obese"))
    )
cus_order_health = c("Exc","V Good", "Good", "Fair", "Poor")

# Begin code.
CVD_cleaned_cus %>% group_by(General_Health,BMI) %>%
  summarise(
    counts = n(), 
    .groups = "keep"
    ) %>%
  ggplot(
    aes(
      x = factor(General_Health,level=cus_order_health),
      y = counts,
      fill = factor(General_Health, level=cus_order_health)
      )
    ) +
  geom_bar(
    position="dodge", 
    stat = "identity"
    ) +
  scale_fill_manual(
    values = c("Exc" = "#0bff00", "V Good" = "#55b100", "Good" = "#859a00", "Fair" = "#d6cb00","Poor" = "#ff0000")
    ) +
  geom_text(
    aes(label = counts),
    size = 2.4,
    vjust = 1
    ) +
# scale = "free_y" gives each BMI scale its own scale.
  facet_wrap(~BMI,scale = "free_y") + 
  labs(x = "BMI Range", 
       y = "Health Count of BMI \n",
       title = "BMI, split by Health", 
       fill = "Health") +
  theme(axis.text.x = element_text(angle = 30,vjust = 0.5)) 
```

```{r General Health grouped by BMI, facet wrap view as total relative frequency}
CVD_cleaned_cus = CVD_cleaned %>% 
  mutate(General_Health = recode(
    General_Health, 
    "Excellent" = "Exc", 
    "Very Good" = "V Good", 
    "Good" = "Good", 
    "Fair" = "Fair", 
    "Poor" = "Poor")
    ) %>%
    mutate(BMI = cut(
    BMI, 
    breaks = c(0,18.5,25,30,35,Inf),
    include.lowest = TRUE,
    right = F,
    labels = c("Underweight", "Healthy", "Overweight", "Obese", "Morbidly Obese"))
    )
cus_order_health = c("Exc","V Good", "Good", "Fair", "Poor")

# Begin code.
CVD_cleaned_cus %>% group_by(General_Health,BMI) %>%
  summarise(
    counts = n(), 
    .groups = "keep"
    ) %>%
  ggplot(
    aes(
      x = factor(General_Health,level=cus_order_health),
      y = counts/sum(counts), 
      fill = factor(General_Health, level=cus_order_health)
      )
    ) +
  geom_bar(
    position="dodge", 
    stat = "identity"
    ) +
  scale_fill_manual(
    values = c("Exc" = "#0bff00", "V Good" = "#55b100", "Good" = "#859a00", "Fair" = "#d6cb00","Poor" = "#ff0000")
    ) +
  geom_text(
    aes(
      label=paste0(
        round(counts/sum(counts)*100,1),"%")
        ),
      vjust=-0.25,
      size=2.5,
      position = position_dodge2(1)
    ) +
  scale_y_continuous(labels=scales::percent,
                    limits=c(0,0.15)) +
  labs(x = "BMI Range", 
       y = "Health Count of BMI \n",
       title = "BMI, split by Health", 
       fill = "Health") +
  facet_wrap(~BMI)
  theme(axis.text.x = element_text(angle = 30,vjust = 0.5)) 
  
# I'm aware this isn't very effective. I did find a conversation online with similar issues and where they tried to have the relative frequencies calculated based on each facet_wrap panel, but it didn't work for me.
```
